---
title: "db_read_node_specific_models"
author: "Joseph Keogh"
date: "11/2/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RPostgreSQL)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(plotly)
library(forecast)
library(broom)
```

## Get data from DB

### Connect to the DB
```{r}
# load credentials
username <- "jgk7uf@va-energy2"
hostname <- "va-energy2.postgres.database.azure.com"
password <- "0Tn5KBFbm2&6bG"
dbname <- "postgres"

# open credentials
db_driver <- dbDriver("PostgreSQL")
db <- dbConnect(db_driver, user=username, password=password, dbname=dbname, host=hostname)

# test connection if returns true the db is connected
print(dbExistsTable(db, "test_table"))
```


### Grab data from DB
```{r}
# select all existing data
response <- dbGetQuery(db, "SELECT * FROM test_table")
```

## Understand Data

### Get basic statistics on data
```{r}
summary(response)
```

The first data is back in June, through beginning of October

3000 Datum

#### The time difference we are looking at
```{r}
as.numeric(max(response$datetime) - min(response$datetime))
```


#### datatypes
```{r}
class(response$datetime)
```

#### what does the rate look like
```{r}
lmp_boxplot <- boxplot(response$total_lmp_rt)
```

#### Can we make the data look better?

##### Remove outliers
```{r}
# remove outliers for boxplot
extreme_high <- lmp_boxplot$stats[5,]

lmp_noxtrm <- filter(response, total_lmp_rt < extreme_high)

boxplot(lmp_noxtrm$total_lmp_rt)

```

Removing extreme values does help the data look better

##### Log transform
```{r}
lmp_log <- mutate(response, total_lmp_rt = log(total_lmp_rt^2+0.00000001))

summary(lmp_log$total_lmp_rt)
boxplot(lmp_log$total_lmp_rt)
```

Performing the log transform does help, but not as much as removing outliers

### Decide on cleaned data
```{r}
extrm_high <- boxplot(response$total_lmp_rt)$stats[5]
extrm_low <- boxplot(response$total_lmp_rt)$stats[1]

response_clean <- response %>%
  filter(total_lmp_rt < extrm_high) %>%
  filter(total_lmp_rt > extrm_low) %>%
  mutate(date = date(datetime)) %>%
  mutate(time = time(datetime)) %>%
  na.omit()

nrow(response)
nrow(response_clean)
summary(response_clean)
boxplot(response_clean$total_lmp_rt)
```

### Separate the data into separate nodes and averaged values for the nodes
```{r}

# separated by node
electric_separate <- response_clean

# averages of the nodes
electric <- response_clean %>%
  group_by(datetime) %>%
  summarize(total_lmp_rt = mean(total_lmp_rt))
```

### Visualize Cleaned data by pnode
```{r}
ggplot(electric_separate, aes(x=datetime, y=total_lmp_rt, group=pnode_name, color=pnode_name)) + 
  geom_line() + 
  labs(title="None Extreme Data", y="Marginal Price", x="Datetime")
```








## End of File