---
title: "DB Reading and Model"
author: "Joseph Keogh"
date: "10/14/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RPostgreSQL)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(plotly)
library(forecast)
library(broom)
```

## Get data from DB

### Connect to the DB
```{r}
# load credentials
username <- "jgk7uf@va-energy2"
hostname <- "va-energy2.postgres.database.azure.com"
password <- "0Tn5KBFbm2&6bG"
dbname <- "postgres"

# open credentials
db_driver <- dbDriver("PostgreSQL")
db <- dbConnect(db_driver, user=username, password=password, dbname=dbname, host=hostname)

# test connection if returns true the db is connected
print(dbExistsTable(db, "test_table"))
```


### Grab data from DB
```{r}
# select all existing data
response <- dbGetQuery(db, "SELECT * FROM test_table")
```

## Understand Data

### Get basic statistics on data
```{r}
summary(response)
```

The first data is back in June, through beginning of October

3000 Datum

#### The time difference we are looking at
```{r}
as.numeric(max(response$datetime) - min(response$datetime))
```


#### datatypes
```{r}
class(response$datetime)
```

#### what does the rate look like
```{r}
lmp_boxplot <- boxplot(response$total_lmp_rt)
```

#### Can we make the data look better?

##### Remove outliers
```{r}
# remove outliers for boxplot
extreme_high <- lmp_boxplot$stats[5,]

lmp_noxtrm <- filter(response, total_lmp_rt < extreme_high)

boxplot(lmp_noxtrm$total_lmp_rt)

```

Removing extreme values does help the data look better

##### Log transform
```{r}
lmp_log <- mutate(response, total_lmp_rt = log(total_lmp_rt^2+0.00000001))

summary(lmp_log$total_lmp_rt)
boxplot(lmp_log$total_lmp_rt)
```

Performing the log transform does help, but not as much as removing outliers

### Decide on cleaned data
```{r}
extrm_high <- boxplot(response$total_lmp_rt)$stats[5]
extrm_low <- boxplot(response$total_lmp_rt)$stats[1]

response_clean <- response %>%
  filter(total_lmp_rt < extrm_high) %>%
  filter(total_lmp_rt > extrm_low) %>%
  mutate(date = date(datetime)) %>%
  mutate(time = time(datetime)) %>%
  na.omit()

nrow(response)
nrow(response_clean)
summary(response_clean)
boxplot(response_clean$total_lmp_rt)
```

### Separate the data into separate nodes and averaged values for the nodes
```{r}

# separated by node
electric_separate <- response_clean

# averages of the nodes
electric <- response_clean %>%
  group_by(datetime) %>%
  summarize(total_lmp_rt = mean(total_lmp_rt))
```


### Visualize the Cleaned data average
```{r}
ggplot(electric, aes(x=datetime, y=total_lmp_rt)) + 
  geom_line() + 
  labs(title="None Extreme Data", y="Marginal Price", x="Datetime")
```

## Create model

### Remove trend

#### Create linear model and see statistical validity
```{r}
lm <- lm(total_lmp_rt ~ datetime, electric)

summary(lm)
```


#### Create linear model and see statistical validity by group
```{r}

# group the data by node
by_node <- group_by(electric_separate, pnode_name)

# create models by node
do(by_node,
   glance(
     lm(total_lmp_rt ~ datetime, data = .)
   ))
```

Model is signficant, see what it looks like on the data

#### Plot the trendline
```{r}
ggplot(electric, aes(x=datetime, y=total_lmp_rt)) +
  geom_line() +
  geom_line(data = fortify(lm), aes(x = electric$datetime, y = .fitted), color="red") +
  labs(title="None Extreme Data with Linear Prediction", y="Marginal Price", x="Datetime")

```

### Use prediction in the future find residuals and save in dataframe
```{r}
electric$lm_res <- lm$residuals
electric$lm_pred <- lm$fitted.values
```

#### Plot the residuals
```{r}
ggplot(electric, aes(x=datetime, y=lm_res)) +
  geom_line() +
  labs(title="Linear Prediction Residuals", y="Marginal Price", x="Datetime")

```

This data looks stationary so we are good to move on

### Remove sinusoidal movement

#### Find any sinusoidal movement
```{r}
# create time series
elec.ts <- ts(electric$lm_res)

# find frequencies of high influence
pgram <- spec.pgram(elec.ts, spans=9, demean=T, log='no')

# sort the frequencies based on influence
sorted.spec <- sort(pgram$spec, decreasing=T, index.return=T)

# convert to periods
sorted.omegas <- pgram$freq[sorted.spec$ix]
sorted.Ts <- 1/pgram$freq[sorted.spec$ix]

# the cutoff for influential
pgram.cutoff <- 5

# the sampling rate per day
print('sampling rate')
nrow(electric)/as.numeric(max(electric$date)-min(electric$date))

# the top periods
print('top periods')
sorted.Ts[1:pgram.cutoff]

# top frequencies
## to double check that this makes sense based on periodogram
print('top frequencies')
sorted.omegas[1:pgram.cutoff]

# visual
pgram.box <- boxplot(sorted.Ts[1:pgram.cutoff], main="Period Boxplot")

# the average influential period
print('mean of top periods')
pgram.box.mean <- pgram.box$stats[3]
print(pgram.box.mean)

# plot top periods
plot(sorted.Ts[1:pgram.cutoff], main = "Top Periods")

### create a model for the seasonality
# assign potential periods to variables
p1 <- sorted.Ts[1]
p2 <- sorted.Ts[2]
p3 <- sorted.Ts[3]
p4 <- sorted.Ts[4]
p5 <- sorted.Ts[4]

# create time variable
time<-c(1:length(elec.ts))

# model
sin_mov <- lm(elec.ts ~ 
                   sin(2*pi*time/p1) + 
                   cos(2*pi*time/p1) +
                   sin(2*pi*time/p2) + 
                   cos(2*pi*time/p2) +
                   sin(2*pi*time/p3) + 
                   cos(2*pi*time/p3) +
                   sin(2*pi*time/p4) + 
                   cos(2*pi*time/p4) +
                   sin(2*pi*time/p5) + 
                   cos(2*pi*time/p5)
                   )

### visualize
ggplot(electric, aes(x=datetime, y=electric$lm_res)) +
  geom_line() +
  geom_line(data = fortify(lm), aes(x = electric$datetime, y = sin_mov$fitted.values), color="red") +
  labs(title="Linear Residuals with Sinusoidal Prediction", y="Marginal Price", x="Datetime")


```

#### Use model to store residuals
```{r}
electric$sin_res <- sin_mov$residuals
electric$sin_pred <- sin_mov$fitted.values
```

##### Plot residuals
```{r}
ggplot(electric, aes(x=datetime, y=sin_res)) +
  geom_line() +
  labs(title="Sinusoidal Prediction Residuals", y="Marginal Price", x="Datetime")

```


If there is still some cyclic movement, we can address this later

### Model Residuals
```{r}
auto <- auto.arima(electric$sin_res, approximation = FALSE)

summary(auto)
```

```{r}
ggplot(electric, aes(x=datetime, y=sin_res)) +
  geom_line() +
  geom_line(aes(x=datetime, y=auto$fitted), colour="red") +
  labs(title="Sinusoidal Prediction Residuals with ARIMA Prediction", y="Marginal Price", x="Datetime")

```

#### Save residual prediction
```{r}
electric$ar_pred <- auto$fitted
```


### Combine all models
```{r}
electric$model_final <- electric$ar_pred + electric$lm_pred + electric$sin_pred
```

### Plot the final model
```{r}
ggplot(electric, aes(x=datetime, y=total_lmp_rt)) +
  geom_line() +
  geom_line(aes(x=datetime, y=model_final), colour="red") +
  labs(title="None Extreme Data with Combined Prediction", y="Marginal Price", x="Datetime")

```

## Model validity and statistics

## End of file
```{r}
"End of file"
```

